Demography and linked selection in wild and domesticated maize

### 2015-08-15 manuscript.tex
almost there

\subsection{Demography of maize domestication}
To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}). 
The most likely model estimates an ancestral $\theta$ of 0.0147 per bp, which translates to an effective population size of , which approximately matches a previous independent estimate \cite{eyre1998}. Assuming a mutation rate of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
The maize population splits from teosinte $\approx 15,000$ generations in the paste with an initial size of only $\approx 5\%$ of ancestral $N_a$. 
Following domestication we estimate considerable gene flow between the populations, estimating $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte.
Our model posits exponential growth in maize following its split from teosinte, and estimates a final modern effective population size of $N_m \approx 370,00$.

@@

\subsection{Post-domestication expansion is apparent in patterns of singleton diversity}
Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thuse younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
We thus repeated our analysis of diversity
 an analysis of genome-wide patterns of diversity using alleles present in only a single individual in the sample. 
These ingleton diversity measures the abundance of alleles that are only observed in one individual among the sample. As a class, singleton alleles depict the most recent patterns of evolution, but
they also have the lowest effect on pairwise diversity. Therefore, unlike
pairwise diversity, patterns of singleton
diversity reflect recent patterns of evolution. Since our sample size
for maize (n=23) was larger than teosinte (n=13), maize singletons were
analyzed directly as well as down-sampled to reflect the sample size of
teosinte. For singletons, our definition of
neutral diversity was defined as diversity at positions at
least 0.02 cM distal from genes. When
evaluating the data in this manner, a very different pattern
emerged compared to what we saw with respect to pairwise diversity. Maize singleton diversity was at least as reduced as teosinte singleton
diversity near genes, but recovered more slowly
(Figure \ref{fig:purify}), implying that in the
recent past maize has been more influenced by purifying selection than has
teosinte.

Together, these findings suggest that demographic history has
a strong influence on the effect of purifying selection. Historically,
teosinte has had a larger population size than maize, and only
recently has maize population size overcome that of teosinte. Since
the efficacy of purifying selection scales with population size,
these results likely reflect changes in $N_e$ more than they reflect
underlying changes in selection pressure.

Similarly, to complement our investigation of pairwise diversity surrounding substitutions, we studied patterns of singleton diversity surrounding synonymous and non-synonymous substitutions in maize (Figure \ref{sFig:singleton}). A nearly identical pattern to that shown in Figure \ref{fig:hardSweeps} was observed. This further demonstrates that even the most recent selection patterns in maize are not dominated by hard sweeps.

>>>>

\subsection{Demography of maize domestication}
To explore whether differences in the efficacy of purifying selection between maize and teosinte can be explained by demographic processes, we estimated the parameters of a simple domestication bottleneck model  (Figure \ref{fig:bottleneck}). 
The most likely model estimates an ancestral population mutation rate of $\theta=0.0147$ per bp, which translates to an effective population size of $N_a \approx 123,000$ individuals given the mutation rate \cite{clark2005}.
The maize population splits from teosinte $\approx 15,000$ generations in the paste with an initial size of only $\approx 5\%$ of ancestral $N_a$. 
Following domestication we estimate considerable gene flow between the populations: $M_{tm} = 1.35$  migrants per generation from teosinte to maize and $M_{mt} = 1.72$ migrants from maize to teosinte. \jri{do you get these numbers by multiplying estimates by Na and modern N_maize? if so we should probably report the proportions not the absolute number of inds.}
After its split from teosinte our model posits exponential population growth in maize, estimating a final modern effective population size of $N_m \approx 370,00$.

@@

\subsection{Population expansion leads to stronger purifying selection in modern maize}
Motivated by the rapid post-domestication expansion of maize evident from our demographic analysis, we investigated whether patterns of diversity at low frequency --- and thus on average younger --- polymorphisms would show similar results in comparisons of maize and teosinte.
We thus repeated our analysis of diversity using alleles present in only a single individual in the sample.
In direct contrast to patterns observed for pairwise nucleotide diversity (Figure \ref{fig:purify}A), diversity of singletons was more strongly reduced in and near genes in maize than in teosinte (Figure \ref{fig:purify}B).  
This pattern remained after downsampling our maize data to account for differences in sample size. 
Taken together, these results suggest that demographic history and population size change have impacted the efficacy of purifying selection during maize evolution.

Patterns of diversity around substitutions, however
